### 2017-05-14 abstract.tex
updated abstract

Backscatter communication enables ultra-low
power wireless transmissions, and is attractive
for networking devices which operate 
on harvested energy. Ambient backscatter
takes the concept further, by leveraging
ambient wireless signals like
television signals, as both the source of power
and carrier signal. Existing state-of-the-art ambient backscatter
systems demonstrate ability to achieve tag-to-tag communication 
under conditions when tags are close to television tower, i.e., with the
strength of TV signals  sufficiently high (-30 dBm).  In this paper, we present our preliminary work
which demonstrates the possibility to backscatter and communicate using signficiantly weaker 
ambient television signals. The key to using weaker broadcast television signals, is to leverage a low-cost
software defined radio receiver (RTLSDR) to receive backscattered transmissions. Our results 
demonstrate that the strength of television signals is strong 
enough in most parts of a mid-sized swedish city for our system
to operate, and communicate, a signficiant improvement over the
state-of-the-art.