### 2017-05-14 paper.tex

\documentclass[conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\newenvironment{itemize*}{\begin{itemize}}{\end{itemize}}
\usepackage[utf8x]{inputenc}
\usepackage{subfigure}
\usepackage{balance}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{caption}
\usepackage{siunitx}
\usepackage{booktabs}

\usepackage{cite}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
 \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
 \usepackage[dvips]{graphicx}
 \DeclareGraphicsExtensions{.eps}
\fi
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{url}

\usepackage[utf8x]{inputenc}

\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
\newcommand{\fakeparagraph}[1]{\vspace{.5mm}\noindent\textbf{#1.}}
\newcommand{\fakepar}[1]{\fakeparagraph{#1}}

\begin{document}

\title{Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio}

\author{\IEEEauthorblockN{Maximilian Stiefel}
\IEEEauthorblockA{Uppsala University\\
maximilian.stiefel.8233@student.uu.se}
\and
\IEEEauthorblockN{Elmar van Rijnswou}
\IEEEauthorblockA{Uppsala University\\
elmar.vanrijnswou.9818@student.uu.se}
\and
\IEEEauthorblockN{Carlos Pérez-Penichet}
\IEEEauthorblockA{Uppsala University\\
carlos.penichet@it.uu.se                              }
\and
\IEEEauthorblockN{Ambuj Varshney}
\IEEEauthorblockA{Uppsala University\\
ambuj.varshney@it.uu.se}
\and
\IEEEauthorblockN{Christian Rohner}
\IEEEauthorblockA{Uppsala University\\
christian.rohner@it.uu.se}
\and
\IEEEauthorblockN{Thiemo Voigt}
\IEEEauthorblockA{Uppsala University\\
and RISE SICS\\
thiemo@sics.se}}

\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
\input{abstract}

\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

\input{introduction}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure}

% Note that the IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command,
% and the \label for the overall figure must come after \caption.
% \hfil is used as a separator to get equal spacing.
% Watch out that the combined width of all the subfigures on a 
% line do not exceed the text width or a line break will occur.
%
%\begin{figure*}[!t]
%\centering
%\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
%\caption{Simulation results for the network.}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat[]), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.
% Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
% labels, the optional argument to \subfloat must be present. If a
% subcaption is not desired, just leave its contents blank,
% e.g., \subfloat[].

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table and, given that table
% captions serve much like titles, are usually capitalized except for words
% such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
% and up, which are usually not capitalized unless they are the first or
% last word of the caption. Table text will default to \footnotesize as
% the IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that the IEEE does not put floats in the very first column
% - or typically anywhere on the first page for that matter. Also,
% in-text middle ("here") positioning is typically not used, but it
% is allowed and encouraged for Computer Society conferences (but
% not Computer Society journals). Most IEEE journals/conferences use
% top floats exclusively. 
% Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the
% \fnbelowfloat command of the stfloats package.

\section{Background}

In this section we introduce some necessary background related to our
work.

%\subsection{Software Defined Radio}
%Traditionally basic radio components like mixers or filters have been realized
%in hardware. Nowadays as soon as one works in the baseband frequency the
%computational resources of a personal computer are sufficient to realize these
%components with signal processing done in software. \cp{We need to expand on
%this introduction to SDR, assuming we need it and not just assume everyone
%knows what that is.}

\subsection{Backscatter Transmissions}
\input{backscatter}

\subsection{RTL2832U}
\label{sub:rtl2832}

\begin{figure}[h]
				\centering
				\includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
				\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q
				demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner
				with integrated LNA \textit{Rafael Micro R820T/2} (right IC).} 
				\label{fig:receiver_arch}
\end{figure}
The \textit{Realtek RTL2832U} is a terrestrial digital video broadcast
(DVB-T) demodulator IC, which is the main part of a wide range of
USB-based DVB-T receiver dongles. One prominent feature of this chip is
that it allows to retrieve the raw I/Q samples via USB. This is
originally intended for the chip to work as a simple DAB/FM software
receiver. Ham radio enthusiasts have combined their efforts to create a
software driver (the \textit{librtlsdr}, cf. \cite{steve-m_librtlsdr})
that allows DVB-T dongles based on the \textit{RTL2832} to be converted
into low-cost wideband SDR receivers. The cost of traditional SDRs has
been in the range of a few hundred to thousands of USD. Albeit much less
powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
can be bought with antenna for less than 10 USD.

An RTL-SDR also contains a tuner that allows the user to select the
received frequency.  In our case the tuner is a \textit{Rafael Micro
R820T/2} which offers a tuning range from 42 to 1002
MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:receiver_arch}, shows a
photograph of the RTL-SDR used in our work.

\subsection{Quadrature Demodulation}

The received signal can be interpreted as
\begin{equation}
	s_{\text{IF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{IF}}(t) \cdot s_{\text{L}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

\section{Design}

In this section we describe the most important aspects of our design.

\subsection{Transmitter}
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.6\columnwidth]{./fig/s11v2}
%\caption{Input reflection coefficient \ensuremath{\Gamma_{in}} also known as \ensuremath{S_{\text{11}}} of the self-made backscattering ground-plane antenna.}
%\label{fig:s11}
%\end{figure}

%A backscattering antenna is the most important part of our backscatter
%transmitter. 
Our backscatter transmitter is designed to be used with the television
signal, which has its center frequency at 626 MHz. To that end, we build
an
antenna using an unetched PCB plate as the ground plane and a wire that acts
as the vertical antenna element.% is soldered to the transmission line. The
%transmission line has been calculated for 626 MHz using an FR4 substrate
%with a thickness of 1.6 mm. The design equations given in
%\cite{pozar_microwave_2011} are implemented in various programs e.g.
%KiCAD, which has been used for solving the design equations. An
%open-stub, attached directly at the juncture where the vertical antenna
%element is attached to the transmission line, is used for fine tuning of
%the input impedance (cf. chapter 5 in \cite{pozar_microwave_2011}).
%Accepting -15 dB as sufficient value for \ensuremath{S_{11}}, one can
%say, that the antenna works in a range from 626 MHz to 645 MHz. For
%backscattering this small bandwidth is not necessarily a disadvantage,
%since everything that is within the operating frequency of the antenna
%is backscattered. With our narrowband antenna the undesired
%backscattering of out-of-band signals is reduced.  With our approach we
%were able to achieve an input impedance at 634 MHz of
%\ensuremath{Z_{in}=(52.2 + j 4.3) \Omega}. At 626 MHz
%\ensuremath{S_{\text{11}}} is still approximately -17 dB.
This antenna works well in the range from 626 to 245 MHz, which has been
measured with a vector network analyzer (VNA).

Besides the antenna, the transmitter consists of an \textit{MSP430} MCU,
that sends data to the receiver. This transmitter controls an RF switch
through an I/O pin. The switch can alternate the impedance connected to
the antenna between matched (50 \ensuremath{\Omega}) and open. The
software now steers this switch with a two 2 MHz rectangular timer
signal to transmit a data 1 or it just leaves the antenna open to
transmit a data 0. Therefore the result is a frequency shift of the
television signal by 2 MHz for a 1 and no shift for a 0. Hence a classical
binary amplitude shift keying is the implemented modulation technique.

\subsection{Receiver}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/receiver_arch}
\caption{Receiver architecture from a system point of view. Signal flow is from left to right.}
\label{fig:receiver_arch}
\end{figure}
The receiver consists of three main modules interconnected by a bus. The
RTL-SDR block is in charge of collecting samples from the SDR-device. The
demodulator turns the sampled signal into a streams of ones and zeroes.
Finally, the Decoder block is in charge of detecting frames. We provide our C++
receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.

%A highly
%sophisticated bus system has been developed to exchange data between the
%different components of the system. 
Figure \ref{fig:receiver_arch} shows the architecture of our receiver, the
signal flows from the left to the right.
%\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
%\textit{libusb}, is used to transfer the data from the TV stick into our
%program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
%down the high frequency to a intermediate frequency (both values are
%software adjustable). Subsequently the data is mixed with sine and
%cosine, low-pass filtered and finally transferred to the digital domain.
The raw I/Q data is available as two 8-bit values (real and imaginary) from the
RTL-SDR. With these two values, the phase and magnitude of the data can be
determined for every sample. The received signal can be represented as:
\begin{equation}
	I+j\,Q = abs(I,Q) \cdot e^{j \times ang(I,Q)} 
\end{equation}

So the first block, entitled RTL2832U, provides the interface to the TV
dongle. It controls the frequency \ensuremath{f_{tuned}}, where the
receiver is listening as well as other interesting parameters e.g. the analog gain.

The demodulator block is responsible for converting the sampled signals
into ones and zeroes. 
%Therefore the real and imaginary values first have
%to be deinterleaved out of the sample message. Now they are represented
%as integer values between 0 and 255 (8 bit). So one converts them to
%floats, whereas 127.5 equals 0. After that the values have to be
%downsampled. Usually we were sampling with 250 kHz. This is still much
%more than we need assuming an Additive White Gaussian Noise (AWGN)
%channel with a \ensuremath{S/N = 10\,dB} (cf.  equation \ref{eq:awgn}).
%So we decided to reduce the sampling frequency to 25 kHz.
%\begin{equation}
%	\label{eq:awgn}
%	C=B \cdot log_2(1+\frac{S}{N})= 25\,kHz \cdot log_2(11) \approx 86.5\,kbit/s
%\end{equation}     
%To do this it is important to use an anti-aliasing filter, as the Shannon-Nyquist theorem has to be satisfied: 
%\begin{equation}
%	\label{eq:nyquist}
%	f_{samp} \geq 2 \cdot f_{a}
%\end{equation}
%\ensuremath{f_{a}} is the highest frequency in the signal. A FIR
%filter naively implemented with floats has been used to achieve this. 10
%kHz is the cut-off frequency of the first filter. After downsampling
%another filter is used to suppress noise. This filter has a cut-off
%frequency of 3 kHz. 
As part of the demodulation process the signal is down-sampled and
low-pass-filtered two times. The last demodulation step is to rectify
the signal (cf. equation \ref{eq:rectify}) and decide with a
software-defined Schmitt trigger whether a sample is a 1 or a 0.  
\begin{equation}
	\label{eq:rectify}
	abs(I,Q) = \sqrt{I^2 +  Q^2}
\end{equation}

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore the code we have written so far subsumes different
simulators for e.g. playing back recorded data. Another handy utility, which
has been written in Octave, is a tool (an oscilloscope) to look at the data,
which is written into files by the demodulator. An example of a recorded data
stream with 25 kS/s can be seen in Figure \ref{fig:transmission}.  
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/transmission}
\caption{Start of a transmission of 101010.. with 1 kbit/s with idle state before. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz.}
\label{fig:transmission}
\end{figure}

\section{Evaluation}
In this section we present our evaluation results. The \textit{RTL-SDR},
combined with intelligent signal processing in \textit{Octave}, are the
main tools we have employed for measurements.

\subsection{Signal Strength Variation in Space}
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_raw}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/626mhz_filtered}
	\end{minipage}
	\caption{Spectrum of the TV signal. Left: Raw spectrum of the TV signal with \ensuremath{f_{\text{center}}=\text{626 Mhz}}. Right: Smoothed spectrum, normalized to maximum average measured in decibel with average shown as red line.}
	\label{fig:tv_record} 
\end{figure}

We have used the \textit{RTL-SDR} to observe the space variations of the
signal strength. \textit{Octave} scripts have been written to aid in
this purpose. These scripts are available under \cite{s3xm3x_RTLSDRSpecAn}. The scripts perform a
frequency sweep through the desired band. The obtained samples are then
transferred from the time domain to the frequency domain using a fast
fourier transformation (FFT).

Because the RTL-SDR has a maximum (stable) sampling rate of 2.4 MS/s, we are
limited to a maximum simultaneous acquisition band of 1.2 MHz. We keep a safety
margin to the maximum capabilities and sample the desired frequency band at 900
kHz intervals and stitch these bands together to form the overall frequency
sweep. This approach is valid only because we are not interested in real-time
data.  With this approach we are able to scan a 20 MHz band in roughly 30
seconds.

During the scanning process the gain is set to 40.2 dB. To get the
signal strength we simply take the average over a 10 MHz band around the
center frequency of the desired TV signal. This approach can be
justified by the fact, that DVB-T is specified up to 10 MHz bandwidth.
And the local TV provider advertises to be transmitting DVB-T 2.
The signal \ensuremath{|H|} is calculated as follows.  
\begin{equation}
	|H| = \Biggl| \sum_{n=0}^N FFT\biggl( \Re\{ s_{\text{band,n}}(t) \} \biggr) \Biggr|
\end{equation}     where \ensuremath{N} is the total number of intervals
accumulated over the entire frequency sweep. The FFT is only carried out over
the real values, which are received from the \textit{RTL-SDR}.
\ensuremath{s_{\text{band,n}}(t)} is the signal in the time domain of one
interval (one subband). There are as many samples taken in one band as needed
for an FFT of size 2048. To finally get \ensuremath{|H_{tv,avg}|} one has to
normalize everything and calculate the power. Before doing that, a simple
smoothing algorithm (FIR) is applied on \ensuremath{|H|}.  
\begin{equation}
	|H_{\text{tv,avg}}| = 20 \cdot log (|H|) - 20 \cdot log(|H_0|)
\end{equation}   
where \ensuremath{|H_0|} is a normalization value measured at a certain point
in space. The result of combining measured signal strength values with the
haversine function to calculate the distance to the point where the maximum
signal has been captured can be observed in Figure \ref{fig:haversine}.
\begin{figure}[h]
	\centering
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/haversine}
	\end{minipage}
	\hfill
	\begin{minipage}{0.49\columnwidth}
	\includegraphics[width=\columnwidth]{./fig/heatmap_626mhz}
	\end{minipage}
	\caption{Left: Signal strength fading against distance relative
to the spot where the maximum signal strength has been captured (TV
towers). 474 MHz and 626 MHz belong to a tower in Uppsala, Vedyxa and
594 MHz belongs to a tower in Uppsala, Rickomberga. Distances have been
calculated with the Haversine formula. Right: Heatmap of the 626 MHz TV
signal in Uppsala.}

\label{fig:haversine}
\end{figure}

\balance

\subsection{Communication Performance}
We tested the communication with different data rates from 1 bit/s up to
1 kbit/s. As can be seen in Figure \ref{fig:transmission} the signal
strength from the backscatter transmitter is rather low. This leads to a
significant amount of quantization noise to appear. This situation can
be improved by increasing the analog gain of the receiver. The
transmission shown in Figure \ref{fig:transmission} was carried out with
the maximum gain (50 dB) available to our receiver.

With a data rate of 1 bit/s, we were able to achieve a range indoors
of a couple of meters. With higher bitrates we
can only communicate over a range of a few decimeters and the 
bit error rate is still approximately 40 \%. The communication
experiment was also carried out outside, where we got similar results.

\section{Discussion}

Our system still has a lot of room for improvement.  Different
backscattering antennas could be tried to find the one with the best
combination of gain, tuning frequency and bandwidth. Moreover the
standard \textit{RTL-SDR} antenna and its coaxial cable are of low
quality.  Hence a custom antenna would be beneficial. Furthermore some
fine tuning can be made when it comes to the filter coefficients of the
different filters in the demodulator chain. Finally  errors can be
further reduced by using one of many available forward error correcting
codes.

To come to a conclusion one can say, that we were able to show, that
spectrum scanning over a huge band with sweeping is possible using the
\textit{RTL-SDR}.  This allowed us to perform a survey of the TV signal
strength in the city. And finally we have shown for the first time, that
backscatter communication is possible with the \textit{RTL-SDR}.

% conference papers do not normally have an appendix

% use section* for acknowledgment
%\section*{Acknowledgment}

%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section
\bibliographystyle{IEEEtran}
\bibliography{paper}

\end{document}