### 2017-05-15 paper.tex

\documentclass[conference]{IEEEtran}
\usepackage{booktabs} % For formal tables
\newenvironment{itemize*}{\begin{itemize}}{\end{itemize}}
\usepackage[utf8x]{inputenc}
\usepackage{subfigure}
\usepackage{balance}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{caption}
\usepackage{siunitx}
\usepackage{booktabs}

\usepackage{cite}

\ifCLASSINFOpdf
 \usepackage[pdftex]{graphicx}
 \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
 \usepackage[dvips]{graphicx}
 \DeclareGraphicsExtensions{.eps}
\fi
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{url}

\usepackage[utf8x]{inputenc}

\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\cp}[1]{\footnote{{\bf Carlos: #1}}}
\newcommand{\fakeparagraph}[1]{\vspace{.5mm}\noindent\textbf{#1.}}
\newcommand{\fakepar}[1]{\fakeparagraph{#1}}

\begin{document}

\title{Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio}

\author{\IEEEauthorblockN{Maximilian Stiefel\IEEEauthorrefmark{1}, Elmar van Rijnswou\IEEEauthorrefmark{1}\\ 
Carlos Pérez-Penichet\IEEEauthorrefmark{1} Ambuj Varshney\IEEEauthorrefmark{1}, 
Christian Rohner\IEEEauthorrefmark{1} and Thiemo Voigt\IEEEauthorrefmark{2}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Uppsala University}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Uppsala Univerisyt and RISE SICS}}

\maketitle

\begin{abstract}
\input{abstract}

\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}

\input{introduction}
\section{Background}

In this section we introduce some necessary background related to our
work.

\subsection{Backscatter Transmissions}
\input{backscatter}

\subsection{RTL2832U}
\label{sub:rtl2832}

\begin{figure}[h]
				\centering
				\includegraphics[width=0.5\columnwidth]{./fig/rtlsdr.jpg}
				\caption{\textit{RTL-SDR} hardware with the DVB-T I/Q
				demodulator \textit{Raeltek RTL2832U} (left IC) and the tuner
				with integrated LNA \textit{Rafael Micro R820T/2} (right IC).} 
				\label{fig:rtl_hw}
\end{figure}
The \textit{Realtek RTL2832U} is a terrestrial digital video broadcast
(DVB-T) demodulator IC, which is the main component in a wide range of
USB-based DVB-T receiver dongles. One prominent feature of this chip is
that it allows to retrieve the raw I/Q samples via USB. This is
originally intended for the chip to work as a simple DAB/FM software
receiver. Ham radio enthusiasts have combined their efforts to create a
software driver (the \textit{librtlsdr}~\cite{steve-m_librtlsdr})
that allows DVB-T dongles based on the \textit{RTL2832} to be converted
into low-cost wideband SDR receivers. The cost of traditional SDRs is
in the range of a few hundred to thousands of USD. Albeit much less
powerful than a typical SDR, a \textit{RTL2832U}-based DVB-T receiver
can be bought with antenna for less than 10 USD.

An RTL-SDR also contains a tuner that allows the user to select the
received frequency.  In our case the tuner is a \textit{Rafael Micro
R820T/2} which offers a tuning range from 42 to 1002
MHz~\cite{rafael_r820t_2011}. Figure \ref{fig:rtl_hw}, shows a
photograph of the RTL-SDR used in our work.

\subsection{Quadrature Demodulation}

The received signal can be interpreted as
\begin{equation}
	s_{\text{RF}}(t)=I(t) \cdot cos(\omega_{0}t) + Q(t) \cdot sin(\omega_{0}t)
\end{equation}     
This is multiplied with a cosine of the carrier frequency from a local oscillator.
\begin{multline}
        s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot cos(\omega_{0}t) \cdot cos(\omega_{0}t)\\+ Q(t) \cdot sin(\omega_{0}t) \cdot cos(\omega_{0}t)
\end{multline}
With \ensuremath{2\,cos(a)cos(b)=cos(a-b)+cos(a+b)} and \ensuremath{2\,sin(a)cos(b)=sin(a+b)+sin(a-b)} follows
\begin{multline}
        2 \cdot s_{\text{RF}}(t) \cdot s_{\text{LO}}(t) = I(t) \cdot \bigl[1+cos(2\omega_{0}t)\bigr]\\+ Q(t) \cdot \bigl[sin(2\omega_0t)+sin(0)\bigr].
\end{multline}
One can see, that the interesting in-phase part (I) in this case is
represented by a DC value after mixing. With a low-pass filter this DC
value can be separated from the undesired rest. An analogous procedure
is done with the quadrature part (Q) when mixing with a sine.

\section{Design}
In this section, we describe the design 
of our system. We first describe the
design of the transmitter, next we describe 
the design of the receiver.

\subsection{Transmitter}

We design our backscatter transmitter to be used with ambient
television signals.  We design the transmitter for ambient
television signal present in the band with  
center frequency of \SI{626}{\mega\hertz}. Crucial to the design
of the transmitter is the implementation of an antenna, as
the antennas are known to be  frequency selective. To this end, we design an
antenna on an unetched board made of FR4 substrate commonly
used to design PCBs. The board acts like the ground plane, at the center
of the board we have a wire acting as a monopole antenna element. 
We have verified using a vector network analyzer (VNA), that the antenna
can operate from frequency range most commonly
used for television broadcast, 
from \SI{245}{\mega\hertz} to \SI{626}{\mega\hertz}.

Backscatter operates by reflecting or absorbing impinging
radio signals on the antenna. To purposefully toggle the antenna
to achieve these states, we  terminate the antenna to a 
RF switch. The RF switch enables us to alternate the impedance connected to
the antenna between matched (50 \ensuremath{\Omega}) and open state.
We control the RF switch through an I/O pin on a low-powered MCU, Texas
instruments MSP430. As discussed earlier, to reduce the effects of
self-interference, we frequency-shift the backscatter transmission
away from the TV signal. Hence, we operate the RF switch using a 
\SI{2}{\mega\hertz} intermediate frequency signal. Therefore the result 
is a frequency shift of the television signal by 2 MHz, when
transmitting a bit 1, and no shift for a 0.  Hence,  we 
leverage amplitude modulation at receiver in our design.
binary amplitude shift keying is the implemented modulation technique.

\subsection{Receiver}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{./fig/receiver_arch}
\vspace{-6mm}
\caption{\emph{Design of the ambient backscatter receiver.} The flow of operation of the
receiver is from the left to right.}
\label{fig:receiver_arch}
\vspace{-6mm}
\end{figure}
The receiver consists of three main modules interconnected by a bus. The
RTL-SDR block is in charge of collecting samples from the SDR-device. The
demodulator turns the sampled signal into a streams of ones and zeroes.
Finally, the Decoder block is in charge of detecting frames. We provide our C++
receiver code publicly available under~\cite{s3xm3x_backscatterBASKReceiver}.

%A highly
%sophisticated bus system has been developed to exchange data between the
%different components of the system. 
Figure \ref{fig:receiver_arch} shows the architecture of our receiver, the
signal flows from the left to the right.
%\textit{Librtlsdr} (cf. \cite{steve-m_librtlsdr}), which is based on
%\textit{libusb}, is used to transfer the data from the TV stick into our
%program. As described in section~\ref{sub:rtl2832}, the RTL2832 mixes
%down the high frequency to a intermediate frequency (both values are
%software adjustable). Subsequently the data is mixed with sine and
%cosine, low-pass filtered and finally transferred to the digital domain.
The raw I/Q data is available as two 8-bit values (real and imaginary) from the
RTL-SDR. With these two values, the phase and magnitude of the data can be
determined for every sample. The received signal can be represented as:
\begin{equation}
	I+j\,Q = abs(I,Q) \cdot e^{j \cdot ang(I,Q)} 
\end{equation}

So the first block, entitled RTL2832U, provides the interface to the TV
dongle. It controls the frequency \ensuremath{f_{tuned}}, where the
receiver is listening as well as other interesting parameters e.g. the analog gain.

The demodulator block is responsible for converting the sampled signals
into ones and zeroes. 
As part of the demodulation process the signal is down-sampled and
low-pass-filtered two times. The last demodulation step is to rectify
the signal (cf. equation \ref{eq:rectify}) and decide with a
software-defined Schmitt trigger whether a sample is a 1 or a 0.  
\begin{equation}
	\label{eq:rectify}
	abs(I,Q) = \sqrt{I^2 +  Q^2}
\end{equation}

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore we have implemented a series of aditional tools that include
different simulators for e.g. playing back recorded data and a tool (an
oscilloscope) to plot the data, which is written into files by the demodulator.
An example of a recorded data stream with 25 kS/s can be seen in Figure
\ref{fig:transmission}.

\begin{figure}[h]
\centering
\vspace{-30mm}

\includegraphics[width=\columnwidth]{./fig/transmission}
\vspace{-30mm}
\caption{Start of a transmission of 101010.. with 1 kbit/s with idle state before. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz.}
\label{fig:transmission}
\vspace{-6mm}

\end{figure}

\section{Evaluation}
\input{eval}

\section{Discussion}

Our system still has a lot of room for improvement.  Different
backscattering antennas could be tried to find the one with the best
combination of gain, tuning frequency and bandwidth. Moreover the
standard \textit{RTL-SDR} antenna and its coaxial cable are of low
quality.  Hence a custom antenna would be beneficial. Furthermore some
fine tuning can be made when it comes to the filter coefficients of the
different filters in the demodulator chain. Finally  errors can be
further reduced by using one of many available forward error correcting
codes.

%To come to a conclusion one can say, that 
We have shown, that
spectrum scanning over a wide band with sweeping is possible using the
\textit{RTL-SDR}.  This allowed us to perform a survey of the TV signal
strength in the city. We found that the signal strength around most of the city
is adequate for backscatter communication using our receiver. And finally we
have shown for the first time, that backscatter communication is possible with
the \textit{RTL-SDR}.

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}