### 2017-05-15 abstract.tex

Ambient backscatter communication enables ultra-low
power wireless transmissions and is hence attractive
for networking devices which operate on harvested energy. 
The data to be sent is modulated on an ambient signal, 
such as TV signals. Tag-to-tag backscatter communication 
requires rather strong ambient signals to achieve acceptable
error rates and range.
In this paper, we present a design and implementation of a
backscatter receiver on a software defined radio plattform 
(RTL-SDR), achieving higher sensitivity than state-of-the-art
backscatter tags and thus enabling wide area ambient backscatter 
communication from low-power tags. 
Our measurements show that the TV signal is strong 
enough in most parts of a mid-sized Swedish city for our
system to operate and communicate.